Fail Faster

### 2025-03-20 paper/paper.tex
Comments on intro (more coming on PDF)

To address this question, we outline two key principles for designing generator libraries that balance efficiency with flexibility.
First, we describe a method of optimizing generator libraries using
\textit{multi-stage programming}, or ``staging,'' which has long been used to
build DSLs without abstraction overhead (``without regret''~\cn{}).
We apply this technique to generators, completely eliminating the run-time overhead of
many common generator abstractions.
Secondly, we demonstrate the impact of different choices of \rand on generator
performance by conducting a controlled comparison.
Together, we refer to this recipe for improving generator libraries as \name{} --- \jwc{something pithy here}.

To demonstrate the effectiveness of our approach, we build two different PBT libraries in two different langauges.
In OCaml, we build \camlname{}, based on \bq{}  --- a
high-quality PBT library in OCaml, authored by software engineers at the trading firm Jane Street \cn{}.
In Scala 3, we build \scalaname{}, based on ScalaCheck --- a very mature PBT library and the de facto choice in Scala \cn{}.
In both languages, our libraries are direct drop-in replacements that are
\textit{100\% semantically equivalent}: that is, given identical random seeds, a generator written with our technique
produces exactly the same sequence of values as the same generator written with the original generator library.
The semantic equivalence between the original generators and \name{} generators allows us to
perform clean, fine-grained comparisons, as differences in generation speed are
solely attributable to the staging and \rand optimizations.

We evaluate the optimizations with a series of case studies in each
generator library, and show that the \name generators run faster than their unstaged equivalents.
Further, we show generators that run faster also find bugs faster
% ability \jwc{The paper is called ``fail faster'' --- we shouldn't say ``bug finding ability'', we should say something like ``how fast you can squash bugs''. This is too passive }
by running these case studies in the Etna framework~\cite{etna}., a
platform that simulates real-world PBT usage\hg{Don't love this phrase, I'll
think more about how we should talk about Etna} by injecting bugs into a system and
measuring how quickly various generators detect them.

@@

Since property-based testing was first introduced in Haskell's QuickCheck
library~\cite{claessenQuickCheckLightweightTool2000}, {\em generators} have been
a key element of the PBT process. While properties are critical for correctly
expressing a developer's intent for a function, the success of the testing
ultimately rests on how well the random inputs to that property exercise the
system under test. QuickCheck's solution to this problem, which has been copied
dozens of frameworks across as many languages, is to provide a domain-specific
language (DSL) for writing random data generators that gives developers complete
control over their distribution of test inputs.

>>>>

To address this question, we outline two key principles for designing
generator libraries that balance efficiency with flexibility.\bcp{We
  just said we want both, not a ``balance'' that trades one off
  against the other...}
First, we describe a method of optimizing generator libraries using
\textit{multi-stage programming}, or ``staging,'' which has long been used to
build DSLs without abstraction overhead (``without regret''~\cn{}).
We apply this technique to generators, completely eliminating the run-time overhead of
many common generator abstractions.
Secondly, we demonstrate the impact of different choices of \rand on generator
performance by conducting a controlled comparison.
Together, we refer to this recipe\bcp{``demonstrating the impact'' is
  not a recipe} for improving generator libraries as \name{} --- \jwc{something pithy here}.

To demonstrate the effectiveness of our approach, we build two different PBT libraries in two different langauges.
In OCaml, we build \camlname{}, based on \bq{}  --- a
high-quality PBT library in OCaml, authored by software engineers at the trading firm Jane Street \cn{}.
In Scala 3, we build \scalaname{}, based on ScalaCheck --- a very
mature PBT library and the {\em de facto} choice\bcp{the wording here
  implies that there are other choices in Scala, some of which might
  be reasonable, and tht base\_quickcheck is not the de facto choice
  in OCaml...} in Scala \cn{}.
In both languages, our libraries are direct drop-in replacements that are
\textit{100\% semantically equivalent}: that is, given identical random seeds, a generator written with our technique
produces exactly the same sequence of values as the same generator written with the original generator library.
The semantic equivalence between the original generators and \name{} generators allows us to
perform clean, fine-grained comparisons, as differences in generation speed are
solely attributable to the staging and \rand optimizations.

We evaluate the optimizations with a series of case studies in each
generator library, and show that the \name generators run faster than their unstaged equivalents.
Further, we show generators that run faster also find bugs faster
% ability \jwc{The paper is called ``fail faster'' --- we shouldn't say ``bug finding ability'', we should say something like ``how fast you can squash bugs''. This is too passive }
by running these case studies in the Etna
framework~\cite{etna}\bcp{``Etna framework'' sounds like more than
  what it is}, a
platform that simulates real-world PBT usage\hg{Don't love this phrase, I'll
think more about how we should talk about Etna} by injecting bugs into a system and
measuring how quickly various generators detect them.

@@

Since property-based testing was first introduced in Haskell's QuickCheck
library~\cite{claessenQuickCheckLightweightTool2000}, {\em generators} have been
a key element of the PBT process. While properties\bcp{this concept
  comes out of nowhere} are critical for correctly
expressing a developer's intent for a function, the success of the testing
ultimately rests on how well the random inputs to that property exercise the
system under test.\bcp{Is this trying to say that generators are more
  important than properties?  This seems like a strange assertion.}
QuickCheck's solution to this problem, which has been copied in
dozens of frameworks across a wide variety of languages, is to provide a domain-specific
language (DSL) for writing random data generators that gives developers complete
control\bcp{that might be overstating the case a bit!} over the
distribution of their test inputs.
